---
title: "The German Tank Problem"
author: "Sage Roy-Burman"
format: html
---

**Introduction:**

During World War II, Allied forces attempted to estimate the total number 
of German tanks produced in order to gain an edge. One of the methods used 
relied on statistical inference using known serial numbers on captured 
tanks. The German tank problem centers on estimating the unknown upper 
bound of the set {1, 2, 3, ... , N}, assuming serial numbers are drawn 
uniformly at random without replacement. This upper bound gives crucial 
information on total tank production. Given a sample of six observed 
serial numbers (43, 86, 171, 36, 177, 222), our goal is to construct three 
different estimators for N, and determine which is the most effective.

**Estimators:**

I decided to compare three estimators: the method of moments estimator 
(MOM), the maximum likelihood estimator (MLE), and the bias corrected MLE, 
which is also the uniformly minimum-variance unbiased estimator (UMVUE). I 
selected the MOM estimator as it is relatively easy to calculate, and is 
also unbiased. I chose the MLE because of its consistency, since as n -\> 
infinity, the MLE converges to the true value. In this scenario, the MLE 
is very intuitive, since it is the largest serial number observed. 
However, since the MLE is downward biased, I also included the UMVUE by 
correcting the MLE for bias. The UMVUE is a very strong contender, since 
it has the lowest variance of all unbiased estimators.

**Simulation Study Design**

To compare these three estimators, I conducted a simulation study in which 
samples were repeatedly drawn without replacement from the set {1, 2, 3, 
... , N}. For each configuration of parameters, I repeated this sampling 
process 5,000 times, calculating the MOM estimator, MLE, and UMVUE for 
every simulated dataset. To look at estimator performance across diverse 
settings, I used four different scenarios combining large and small values 
of N and n. Specifically, I considered N = 250, n = 6; N = 2000, n = 6; N 
= 250, n = 30; and N = 2000, n = 30. This mix of limited and informed 
sampling conditions allows for better comparison of estimator behavior. 
For each estimator and scenario, I computed four performance metrics: the 
mean estimate across simulations, the bias, the standard deviation, and 
the mean squared error.

**Simulation Study Results**

```{r, include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
# Load necessary packages
```

```{r, echo = FALSE}
single_trial <- function(N, n) {
  x <- sample(1:N, n, FALSE)
  xbar <- mean(x)
  xmax <- max(x)
  c(
    MLE = xmax,
    MOM = 2*xbar - 1,
    UMVUE = ((n + 1)/n) * xmax - 1
  )
}
# This function performs a single simulation of the problem, calculating 
# the three estimators for the one sim.
```

```{r, echo = FALSE}
run_sim <- function(N, n, R = 5000) {
  sims <- replicate(R, single_trial(N, n))
  sims <- t(sims)
  
  data.frame(
    Estimator = colnames(sims),
    Mean = colMeans(sims),
    Bias = colMeans(sims) - N,
    SD = apply(sims, 2, sd),
    MSE = colMeans((sims-N)^2)
  )
}
# This function runs the simulation study for a given population size and 
# sample size. It repeats the sampling process 5,000 times, computes each 
# estimator for each simulated dataset, and calculates mean, bias, SD, and MSE.
```

```{r, echo = FALSE}
small_N_small_n <- run_sim(250, 6)
large_N_small_n <- run_sim(2000, 6)
small_N_large_n <- run_sim(250, 30)
large_N_large_n <- run_sim(2000, 30)
# This block applies the simulation function to the four scenarios.
```

```{r, echo = FALSE}
all_results <- bind_rows(
  small_N_small_n,
  large_N_small_n,
  small_N_large_n,
  large_N_large_n
) %>%
  select(Estimator, Mean, Bias, SD, MSE)

kable(all_results, digits = 2, row.names = FALSE,
      caption = "Mean, bias, empirical standard deviation, and MSE for each estimator") %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  pack_rows("N = 250, n = 6", 1, 3) %>%
  pack_rows("N = 2000, n = 6", 4, 6) %>%
  pack_rows("N = 250, n = 30", 7, 9) %>%
  pack_rows("N = 2000, n = 30", 10, 12)
# This table summarizes the performance of each estimator across scenarios.
```

The simulation results reveal many clear estimator patterns across all 
scenarios. First, the MLE exhibits downward bias in every setting. 
Although the variability of the MLE is low (from the standard deviation), 
the MSE is consistently relatively large. For the MOM estimator, the bias 
is extremely low, as the estimator itself is unbiased. However, it has the 
highest standard deviation, and therefore variability, of all estimators 
in all scenarios, resulting in a very high MSE. Overall, the UMVUE 
performs the best of all estimators, which is unsurprising. Its 
combination of low variability and lack of bias results in the lowest MSE 
over all settings. These effects are even more prominent when the sample 
size is lower, as bias and variability are more pronounced.

```{r, echo = FALSE}
library(ggplot2)

ggplot(bind_rows(
  mutate(small_N_small_n, Scenario = "N = 250, n = 6"),
  mutate(large_N_small_n, Scenario = "N = 2000, n = 6"),
  mutate(small_N_large_n, Scenario = "N = 250, n = 30"),
  mutate(large_N_large_n, Scenario = "N = 2000, n = 30")
), aes(x = Estimator, y = MSE)) +
  geom_col() +
  facet_wrap(~Scenario, scales = "free_y") +
  labs(
    title = "Mean Squared Error of Different Estimators",
    x = "Estimator",
    y = "MSE"
  )
# This code plots the MSE of different estimators across scenarios
```

**Figure 1: MSE of Different Estimators Across Scenarios** shows that the 
UMVUE consistently has the lowest MSE, with the MLE following and the MOM 
estimator having consistently large MSE.

```{r, echo = FALSE}
library(tidyverse)

sim_data <- replicate(5000, single_trial(n = 6, N = 250))
sim_data <- t(sim_data)
sim_data <- as.data.frame(sim_data)

sim_data_long <- sim_data %>%
  pivot_longer(cols = everything(),
               names_to = "Estimator",
               values_to = "Estimate")
# This code generates 5,000 simulated samples for a fixed scenario, and 
# computes the three estimators for each sample. It is then reshaped into 
# long format, where each row corresponds to a single estimate and its 
# associated estimator. This format is used for easier vizualization.
```

```{r, echo = FALSE}
ggplot(sim_data_long, aes(x = Estimator, y = Estimate, fill = Estimator)) +
  geom_boxplot(alpha = 0.7) +
  geom_hline(yintercept = 250, color = "red", linetype = "dashed") +
  labs(
    title = "Distribution of Estimators for N (N = 250, n = 6)",
    y = "Estimated N"
  )
# This plot shows the distribution of estimators across simulated samples for a specific scenario.
```

**Figure 2: Distribution of Estimators for N = 250, n = 6** illustrates 
how the MOM estimator shows extreme variability, whereas the MLE and UMVUE 
have a tighter spread. Additionally, the downward bias of the MLE is 
apparent, and the MOM and UMVUE appear unbiased.

**Application to Observed Data**

Using the observed serial numbers, {43, 86, 171, 36, 177, 222}, we are 
able to calculate the three estimators. 
$$
\hat{N}_{\text{MLE}} = 222
$$

$$
\hat{N}_{\text{MOM}} = 244
$$

$$
\hat{N}_{\text{UMVUE}} = 258
$$

Since our simulation results showed the UMVUE performed better in terms of 
bias and mean squared error, we should prefer the UMVUE estimate.

**Conclusion**

The German tank problem highlights the importance of the balance between 
bias and variance in selecting the proper estimator. For the maximum 
likelihood estimator, it exhibited low variance, but its downward bias 
resulted in consistent underestimation in the simulations. The method of 
moments estimator is unbiased, but had extremely high variability. The 
UMVUE combines the strengths of both, giving low bias while also limiting 
variance. As a result of the simulation evidence, the UMVUE is the clear 
preferred estimator. Using the UMVUE, I approximate that N = 258 tanks 
were produced.
